### 2019-03-19 supp.tex
Fix up SI

\documentclass[aps,rmp,onecolumn,
			   groupedaddress,superscriptaddress,
			   amsfonts,amssymb,amsmath,
			   citeautoscript,bibnotes,
			   a4paper]{revtex4}

%-----PACKAGES-----
\usepackage[pdftex]{hyperref}
\hypersetup{colorlinks,
	linkcolor={blue!75!black!80!yellow},
	citecolor={blue!75!black!80!yellow},
	urlcolor={blue!75!black!80!yellow},
	pdfstartview=FitH}
\usepackage[top=1in, bottom=1.5in, left=1.25in, right=1.25in]{geometry} %Change margins
\usepackage{graphicx}
\usepackage{xspace}
\newcommand\bmmax{0} % the bm package takes up a lot of alphabets internally, making it impossible to load more font packages (like sansmath); reduce the number of preloaded alphabets to avoid this
\usepackage{bm}
\usepackage{braket}
\usepackage{xcolor}
\usepackage[UKenglish]{babel}

\usepackage{siunitx}
\sisetup{range-phrase =\text{\,--\,},
	list-units   =single,
	range-units  =single,
	list-pair-separator = {\ \text{and}\ },
	list-separator = {,\ \linebreak[0]},
	list-final-separator = {,\ \linebreak[0]\text{and}\ },
	detect-all = true}
\DeclareSIUnit[number-unit-product = ]\percent{\char`\%} % remove space before percentage sign

%-----APPPEND 'S' TO REFERENCES OF ALL KINDS-----
\renewcommand{\theequation}{S\arabic{equation}} %Equations
\renewcommand{\thefigure}{S\arabic{figure}} %Figures
\renewcommand{\bibnumfmt}[1]{[S#1]} %Citations
\renewcommand{\citenumfont}[1]{S#1}
\renewcommand{\thetable}{Supplementary Table \arabic{table}}%
\renewcommand*\descriptionlabel[1]{\hspace\labelsep\normalfont\itshape #1\ \hfil} %Change appearance of description lists
\renewcommand{\thesection}{S\arabic{section}}
\renewcommand{\thesubsection}{\Alph{subsection}}
\renewcommand{\thesubsubsection}{\roman{subsubsection}}

%-----LOADING RELEVANT FONTS [Times Roman]-----
\usepackage{txfonts}  %Times Roman fonts
\usepackage{txfontsb} %Addition for txfonts, including old style numerals and greek

%-----REDEFINE THE RMP STYLE OF AUTHORS AND AFFILIATIONS-----
\makeatletter
\def\frontmatter@authorformat{%
	\preprintsty@sw{\vskip0.5pc\relax}{}%
	\@tempskipa\@flushglue
	\@flushglue\z@ plus50\p@\relax
	\raggedright\advance\leftskip.25in\relax
	\@flushglue\@tempskipa
	\parskip\z@skip
}%
%
\def\frontmatter@affiliationfont{% Helvetica 9/10.2
	\small\slshape\selectfont\baselineskip10.5\p@\relax
	\@tempskipa\@flushglue
	\@flushglue\z@ plus50\p@\relax
	\raggedright\advance\leftskip.25in\relax
	\@flushglue\@tempskipa
}
%	
\def\paragraph{%
	\@startsection
	{paragraph}%
	{4}%
	{\parindent}%
	{\z@}%
	{-1em}%
	{\normalfont\small\itshape\sansmath\textsf}%%Change so that paragraphs are also in sans serif font
}%
\renewcommand*\email[1][]{\begingroup\sanitize@url\@email{#1}} %Remove the 'Electronic adress: ' from RMP style
\makeatother

\setcitestyle{numbers,square,sort&compress} %We don't want the rmp citation style, but standard style square bracket, numbers style instead

%----- Changes to caption -----
% These hacks are specific to revtex4 (revtex4-1 solution doesn't work)
\usepackage[EULERGREEK]{sansmath}
\makeatletter 
% set figure caption title to bold sans serif... 
\renewcommand*{\fnum@figure}{{\sffamily\bfseries{}\figurename~\thefigure{}.}}
% set figure caption contents to sans serif
\long\def\@makecaption#1#2{%
	\vskip\abovecaptionskip
	\vbox{%
		\flushing%
		\sisetup{math-sf=\textsf}%
		\small\sffamily\sansmath%
		\noindent
		#1\nobreak\hskip.5em plus.2em\ignorespaces#2\par
	}%
	\vskip\belowcaptionskip
}%
\makeatother

%----- Spacing and binary relations/operations -----
\thickmuskip=5mu plus 2mu minus 1mu  %binary relations (default, 5mu plus 5mu)
\medmuskip=4mu plus 2mu minus 2mu    %binary operations (default, 4mu plus 2mu minus 4mu)
\frenchspacing %Ensure that revTeX does not do "double spaces" after punctuation

%-----COMMANDS-----
\def\changemargin#1#2{\list{}{\rightmargin#2\leftmargin#1}\item[]}
\let\endchangemargin=\endlist 
\newcommand{\appropto}{\mathrel{\vcenter{
			\offinterlineskip\halign{\hfil$##$\cr
				\propto\cr\noalign{\kern2pt}\sim\cr\noalign{\kern-2pt}}}}}

% hyphenation rules
\hyphenation{}

% --- COMMANDS ---
\renewcommand{\Im}{\operatorname{Im}}
\renewcommand{\Re}{\operatorname{Re}}
\newcommand{\sgn}{\operatorname{sgn}}

\newcommand{\iu}{\mathrm{i}}
\newcommand{\e}{\mathrm{e}}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\unitvec}[1]{\ensuremath{\hat{\mathbf{#1}}}}
\newcommand{\sub}[1]{\ensuremath{_{\textrm{#1}}}} %Upright multi-character subscript
\newcommand{\super}[1]{\ensuremath{^{\textrm{#1}}}} %Upright multi-character superscript

%abbreviations
\newcommand{\ie}{i.e.\@\xspace} %Gobble-spaces of the "small" type (small is ensured by adding \@)
\newcommand{\cf}{cf.\@\xspace}
\newcommand{\eg}{e.g.\@\xspace}
\newcommand{\etc}{etc.\@\xspace}

% ---- Author affiliations -----
\newcommand{\HarvardSEAS}{John A. Paulson School of Engineering and Applied Sciences, Harvard University, Cambridge, MA, 02139, USA}
\newcommand{\MITPhy}{Department of Physics, Massachusetts Institute of Technology, Cambridge, MA, 02139, USA}

% --- DOCUMENT ---
\begin{document}

% --- TITLE & AUTHORSHIP ---
\title{SUPPORTING INFORMATION\\Phonon polaritonics in two-dimensional materials}

\author{Nicholas Rivera}
\email{nrivera@seas.harvard.edu}
\affiliation{\HarvardSEAS}\affiliation{\MITPhy}
%
\author{Thomas Christensen}
\affiliation{\HarvardSEAS}\affiliation{\MITPhy}
%
\author{Prineha Narang}
\email{prineha@seas.harvard.edu}
\affiliation{\HarvardSEAS}

% --- ABSTRACT ---

\maketitle

\vspace*{4ex}
\begin{figure*}[h]
    \includegraphics[width=0.725\linewidth]{enhancement.png}
    \caption{%
        \textbf{Spontaneous emission enhancement in hBN thin films and monolayer.} 
        Spontaneous emission enhancement for a dipole emitter \SI{5}{\nm} away from the surface of thin films of bulk hBN as a function of emission frequency for film thicknesses of \SIlist{100; 10; 1; 0.33}{\nm}, as well as monolayer hBN (dispersion relation and conductivity given in the main text). 
        In all cases, the hBN is assumed surrounded by air. 
        The bulk TO frequency is \SI{1387}{\per\cm}.
        \label{fig:s1}
        }
\end{figure*}

\mbox{}
\vfill

\mbox{}

\renewcommand\refname{} % we don't have any references this time around, so suppress printing of 'References' section head
%\bibliographystyle{apsrev4-1}
%\bibliography{references}

\end{document}